\vspace{1.0cm} \Large \bf Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''

### 2016-04-28 mSystems/mSys_results.tex
Some work in mSys Results section (v0.0.9)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex
	modified:   results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf
	modified:   results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf
	modified:   results/hist/IBS_h_A_amplicons_family_hist2D.pdf
	modified:   results/hist/antibiotics_h_patD_pre_AbsFreqPlot.pdf

\subsubsection*{Fit Plots}
For each and every dataset included in the study, an unweighted fit (see Material and Methods section for details) and a X-weighted fit (detailed in Material and Methods subsection) have been calculated for standard deviations versus the mean values for each bacterial genus monitored in time. Figure \ref{fig:main1} showed the X-weighted fit for samples from a healthy subject (patient A, top) and from a subject diagnosed with irritable bowel syndrome (patient P2, bottom) studied in our lab\cite{IBS}, while Figure \ref{fig:fits} shows the corresponding unweighted fits. Additionally, for the unweighted fit, a complete residues analysis was performed, and a 4-in-1 figure was generated as shown in Figure \ref{fig:unwRes}, corresponding to patient A (top plot in Figure \ref{fig:main1} and Figure \ref{fig:fits}). Among other tests, it allows to check for normality and homoscedasticity of the residues and, therefore, to further evaluate the goodness of the fit.

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P1_metatranscriptomes_family_stdVSmean_LLR_LOG.pdf}
	\caption{Log plots of unweighted fits corresponding to the datasets shown in Figure \ref{fig:main1}}
	\label{fig:fits}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
	\caption{Residues analysis plot corresponding to the unweighted fit for patient A of the IBS study\cite{IBS}. The top-left subplot is a simple residues plot. The top-right subplot is a Normal quantiles plot with linear fitting (value of coefficient of determination is provided). The bottom-left subplot shows an accumulated ESS (Explained Sum of Squares) plot. Finally, the bottom-right subplot is a residues Normal histogram plot. This set of subplots allows to check for normality and homoscedasticity of the residues.}
	\label{fig:unwRes}
\end{figure}

\subsubsection*{Histogram Plots} 
\CC\ generates two related histogram plots, absolute frequencies plot and zero relative frequency plot. The former is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem. In Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}. On the other hand, we could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found. Attending to all taxa, we can plot the ZRF histogram, which then lies on the horizontal axis of the plot. The vertical axis shows the number of taxa, so the height of a bar represents the amount of taxa that have determinate ZRF. In this respect, the bar over $0.0$ counts the quantity of taxa that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the total of taxa that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the software. Figure \ref{fig:histZRF} shows this plot for the healthy patient A of the IBS study performed in our lab\cite{IBS}. There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once. So, this plot is clearly useful to notice how the ``core'' is distributed.

\begin{figure}		
	\centering		
 	\includegraphics[width=0.8\textwidth]{results/hist/antibiotics_h_patD_pre_AbsFreqPlot}		
 	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}. We can see that there are no \emph{outlayers} among the total taxa sum for the time series points, so all of them were considered for further analysis by \CC\ software}		
 	\label{fig:histAFP}		
\end{figure}

>>>>

\subsubsection*{Fit Plots}
For each and every dataset included in the study, an unweighted fit and a X-weighted fit (see Material and Methods section for details on both fits) have been calculated for standard deviations versus the mean values for each bacterial genus monitored in time. Figure \ref{fig:main1} showed the X-weighted fit for samples from a healthy subject (patient A, top) and from a subject diagnosed with irritable bowel syndrome (patient P2, bottom) studied in our lab\cite{IBS}, while Figure \ref{fig:fits} shows the corresponding unweighted fits. Both for the X-weighted and the unweighted fit, the IBS patient shows significative higher values of $V$ and $\beta$ than the healthy subject.

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P1_metatranscriptomes_family_stdVSmean_LLR_LOG.pdf}
	\caption{Log plots of unweighted fits corresponding to the datasets shown in Figure \ref{fig:main1}, a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom) studied in our lab \cite{IBS}}
	\label{fig:fits}
\end{figure}

Additionally, for the unweighted fit, a complete residues analysis was performed, and a 4-in-1 figure was generated as shown in Figure \ref{fig:unwRes}, corresponding to patient A (top plot in Figure \ref{fig:main1} and Figure \ref{fig:fits}). This allows to check for normality and homoscedasticity of the residues and, therefore, to further assess the goodness of the model's fit to the data.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
	\caption{Residues analysis plot corresponding to the unweighted fit for patient A of the IBS study\cite{IBS}. The top-left subplot is a simple residues plot. The top-right subplot is a Normal quantiles plot with linear fitting (value of coefficient of determination is provided). The bottom-left subplot shows an accumulated ESS (Explained Sum of Squares) plot. Finally, the bottom-right subplot is a residues Normal histogram plot. This set of subplots allows to check for normality and homoscedasticity of the residues and, therefore, to further assess the goodness of the model's fit to the real data.}
	\label{fig:unwRes}
\end{figure}

\subsubsection*{Histogram Plots} 
The \CC\ software generates two related histogram plots: absolute frequencies plot and zero relative frequency plot. The former is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem. In Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}. We can see that there are no \emph{outlayers} among the total taxa sum for the sample points of this time series, so all of them were considered for further analysis by \CC.

On the other hand, we could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found. Attending to all taxa, we can plot the ZRF histogram, which then lies on the horizontal axis of the plot. The vertical axis shows the number of taxa, so the height of a bar represents the amount of taxa that have determinate ZRF. In this respect, the bar over $0.0$ counts the quantity of taxa that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the total of taxa that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the software. Figure \ref{fig:histZRF} shows this plot for the healthy patient A of the IBS study performed in our lab\cite{IBS}. There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once. So, this plot is clearly useful to notice how the ``core'' is distributed.

\begin{figure}		
	\centering		
 	\includegraphics[width=0.8\textwidth]{results/hist/antibiotics_h_patD_pre_AbsFreqPlot}		
 	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}. There are no \emph{outlayers} among the total taxa sum for the time series sample points, so all of them were considered for further analysis by \CC\ software}		
 	\label{fig:histAFP}		
\end{figure}
